---
title: "BREXIT - Synthetic Controls"
output: github_document
date: "2026-04-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
source(here::here("worldbankdata","romania_synthetic.R"))
```

```{r, libraries, echo=FALSE,include=FALSE,warning=FALSE}
library(AER)
library(patchwork)
library(tidyverse)
library(dplyr)
library(purrr)
library(gridExtra)
library(readr)
library("MatchIt")
library("cobalt")
library("matrixStats")
library(ivreg)
library("rdrobust")
library("estimatr")
library("Synth") 
library("tidysynth")  
library("jtools")
library(tidyr)
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(tidysynth)
```


# Introduction

BREXIT synthetic control analysis.  


## Synthetic UK after 2016 referendum




## Control and variable weights

```{r, echo = FALSE, warning = FALSE}
# Weighting of units and variables
gdp_out2 %>% plot_weights()
```

It looks like Ukraine was the most suitable in constructing a synthetic Romania pre-2007. Turkey, Indonesia and South Korea also contribute a lot. The variable given the most weight is average gdp per capita.  

## Synthetic and treated comparison

```{r, echo=FALSE,warning=FALSE, message=FALSE}
# Comparability of synthetic control to treated unit
gdp_out2 %>% grab_balance_table()
```

The algorithm's done a good job - mostly - in creating a synthetic Romania. Average GDP, imports and average government consumption,  are especially similar. The only major flaw is that real Romania has negative population growth, which the synthetic doesn't. 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
#graph comparing paths
gdp_out2 %>% plot_trends()
```

There is a clear jump in real GDP per capita for Romania after accession to the EU compared to the synthetic. 

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# Graph of difference between the two
gdp_out2 %>% plot_differences()
```

This can also be seen by plotting the differences. Before the accession, the difference moved around a mean of 0. After, the gap fairly consistently grows. 

# Excluding COVID Years
I also try excluding 2020 and 2021 in case COVID was messing with the inference. 

```{r, echo=FALSE,warning=FALSE, message=FALSE}
# Comparability of synthetic control to treated unit
gdp_out3 %>% grab_balance_table()
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
#graph comparing paths
gdp_out3 %>% plot_trends()
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# Graph of difference between the two
gdp_out3 %>% plot_differences()
```