---
title: "Factors Influencing Marriage"
output: github_document
date: "2026-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AER)
library(plm)
library(lmtest)
library(dplyr)
library(tidyr)
library(stargazer)
library(webshot2)
library(htmltools)
library(webshot2)
library(ggplot2)
library(tidyverse)
library(kableExtra)
```

## Marriage

I perform a logistic regression on the CPS1985 data - cross-sectional data which comes from the May 1985 current population survey by the US Census Bureau. 
```{r source-script, include = FALSE}
source(here::here("CPS1985.R"))
```

Here's some plots of the ditribution of various statistics:

```{r plot-edu-wage, fig.show='hold',out.width = "50%",fig.height=2, fig.width=3, echo=FALSE}
ggplot(df, aes(x = education)) +
  geom_histogram(fill = "#69b3a2",binwidth = 1) +
  labs(x = "years of education", y = "count") +
  ggtitle("Years of education \n distribution") +
  theme(plot.title = element_text(size = 9, face = "bold"))

ggplot(df, aes(x = wage)) +
  geom_histogram(binwidth = 1, fill = "#69b3a2") +
  labs(x = "Wage (dollars/hour)", y = "count") +
  ggtitle("Wage distribution") +
  theme(plot.title = element_text(size = 9, face = "bold"))

```


```{r plot-age-ethnicity, fig.show='hold', out.width = "50%",fig.height=2, fig.width=3, echo=FALSE}
ggplot(df, aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "#69b3a2") +
  labs(x = "Age", y = "count") +
  ggtitle("Age distribution") +
  theme(plot.title = element_text(size = 9, face = "bold"))

ggplot(eth, aes(x = ethnicity, y = count)) +
  geom_bar(stat = "identity",fill = "#69b3a2") +
  labs(x = "Ethnicity", y = "count") +
  ggtitle("Ethnicity distribution") +
  theme(plot.title = element_text(size = 9, face = "bold"))
  
```

The logit equation used is 
```{r, include=TRUE, echo=TRUE}
logit <- glm(married ~ wage + age + ethnicity + education + gender, 
             data = df,
             family = binomial(link = "logit"))
```
.      

First, I predict difference in probability of being married where all characteristics between two people are the same - set to the mean - except one person has the mean number of years' education, and the other has the highest, at **`r max(df$education)` years**. The difference comes out as **`r round(edu_predic,4)`.** This implies that going from **`r round(mean(df$education),0)` years** of education to **`r max(df$education)`** decreases one's probability of being married by **`r 100 * round(edu_predic,4)`** percentage points. 

Next I compute how probability of being married changes with different paris of wage changes. The table below shows the 8 different wage pairs and the corresponding increase/decrease in probability of being married. 


**Predicted probability change in being married**
```{r preds, fig.cap="Predictions", echo=FALSE}
knitr::kable(results_df)
```

The table above demonstrates that, evidently, the higher the hourly wage, the higher the chance of being married. 