---
title: "Factors Influencing Marriage"
output: github_document
date: "2026-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AER)
library(plm)
library(lmtest)
library(dplyr)
library(tidyr)
library(stargazer)
library(webshot2)
library(htmltools)
library(webshot2)
library(ggplot2)
library(tidyverse)
```

## Marriage

I perform a logistic regression on the CPS1985 data - cross-sectional data which comes from the May 1985 current population survey by the US Census Bureau. 
```{r source-script, include = FALSE}
source(here::here("CPS1985.R"))
```


```{r plot, fig.height=2, fig.width=3, echo=FALSE}
ggplot(df, aes(x = education)) +
  geom_histogram(fill = "#69b3a2",binwidth = 1) +
  labs(x = "years of education", y = "count") +
  theme(plot.title = element_text(size = 9, face = "bold"))
```

Here we see the distribution of number of years of education, the mean value being `r round(mean(df$education),2)`.

The logit equation used is logit <- glm(married ~ wage + age + ethnicity + education + gender, 
             data = df,
             family = binomial(link = "logit")).
             
First, I predict difference in probability of being married where all characteristics between two people are the same - set to the mean - except one person has the mean number of years' education, and the other has the highest, at **`r max(df$education)` years**. The difference comes out as **`r round(edu_predic,4)`.** This implies that going from **`r round(mean(df$education),0)` years** of education to **`r max(df$education)`** decreases one's probability of being married by **`r 100 * round(edu_predic,4)`** percentage points. 
