---
title: "Money Supply"
output: github_document
date: "2026-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(scales)
library(AER)
library(plm)
library(lmtest)
library(dplyr)
library(tidyr)
library(stargazer)
library(webshot2)
library(htmltools)
library(webshot2)
library(ggplot2)
library(tidyverse)
library(kableExtra)
```

```{r, include=FALSE}
source(here::here("money_supply.R"))
```

## Money Supply

I want to see whether year-on-year money supply growth affects year-on-year inflation. 

Using data from <https://www.bankofengland.co.uk/statistics/Tables>, Table A2.2.1 - Components of M4, I plot the change in UK money supply. Since most of the money supply is created when commerical banks create deposits, I use total retail deposits for households and private non-financial corporations. Wholesale deposits and repos are excluded, as well as notes and coin. 

The data runs from `r min(df_growth$Date)` to `r max(df_growth$Date)`. I plot total money supply in sterling millions and year-on-year money supply growth.

```{r plots, fig.height=2, fig.width=7, echo=FALSE, warning = FALSE}
ggplot(df, aes(x = Date, y = M4)) + 
  geom_line(color = "yellow", linewidth = 0.5) +
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Growth") +
  ggtitle("Money Supply\nsterling millions 1982 - 2026") +
  theme(plot.title = element_text(size = 12 , face = "bold", colour = "white"))+
  theme(panel.background = element_rect("black")) +
  theme(plot.background = element_rect("black"))
```

The money supply as of `r format(max(df$Date))` is 
£`r scales::comma(1e6 * (df %>% filter(Date == max(Date)) %>% slice(1) %>% pull(M4)))`, i.e. £`r scales::comma(round((1e6 * (df %>% filter(Date == max(Date)) %>% slice(1) %>% pull(M4))) / 1e12,3))` trillion.


```{r more-plots, fig.height=2, fig.width=7, fig.show='hold', echo=FALSE, out.width="50%", echo=FALSE, warning=FALSE}
ggplot(df_growth, aes(x = Date, y = money_growth_yoy)) + 
  geom_line(color = "yellow", linewidth = 0.5) +
  labs(x = "Year", x = "Growth") +
  scale_y_continuous(labels = comma) +
  ggtitle("% M4 growth\nYear on year") +
  theme(plot.title = element_text(size = 12 , face = "bold", colour = "white"))+
  theme(panel.background = element_rect("black")) +
  theme(plot.background = element_rect("black"))

```


For year on year growth rates, the highest growth occurred on `r df_growth %>% filter(money_growth_yoy  == max(money_growth_yoy , na.rm = TRUE)) %>% slice(1) %>% pull(Date)`, with a growth rate of **`r paste0(round(df_growth %>% filter(money_growth_yoy   == max(money_growth_yoy , na.rm = TRUE)) %>% slice(1) %>% pull(money_growth_yoy ),2), "%")`**.

## Inflation

I plot year on year inflation data from <https://www.ons.gov.uk/economy/inflationandpriceindices/timeseries/l55o/mm23>. Compare it to money supply growth.

```{r inflation-plot, fig.height=2, fig.width=7, echo=FALSE}
ggplot(data = inf, aes(x = Date)) +
  geom_line(aes(y = inflation, color = "inflation"), linewidth = 0.8) +
  geom_line(aes(y = money_growth_yoy, color = "money_growth_yoy"), linewidth = 0.8) +
  scale_color_manual(
    name = "Legend",
    values = c("inflation" = "red", "money_growth_yoy" = "blue"),
    labels = c("Inflation", "Money growth")
  ) +
  labs(x = "Date", y = "Year") +
  ggtitle("Money Supply\nand Inflation 1989 - 2026") +
  theme(axis.title.x = element_text(color = "white")) +
  theme(axis.title.y = element_text(color = "white")) +
  theme(plot.title = element_text(color = "white", face = "bold", size = 12)) +
  theme(plot.background = element_rect("black")) +
  theme(panel.background = element_rect("black"))
```

Between `r min(inf$Date)` and `r max(inf$Date)`, the highest recorded rate of inflation was **`r paste0(inf %>% filter(inflation == max(inflation)) %>% slice(1) %>%  pull(inflation),"%")`** which occurred in `r inf %>% filter(inflation == max(inflation)) %>% slice(1) %>% pull(Date)`

## VAR Model

I run a VAR model on the data. I first did a simple dynlm but wasn't happy with the results. The benefit of VAR is that all variables are endogenous, so money growth is allowed to affect inflation and vice-versa. This is probably more realistic since inflation can cause money growth movements via central bank behaviour, and money supply growth can affect inflation behaviour too. 

```{r, include=FALSE}
#VAR
var_data <- cbind(inflationn = inf$inflation, 
                  m4 = inf$money_growth_yoy)
```

```{r, include = FALSE}
var_data <- ts(var_data, start = start(inf$inflation), frequency = 12)
```

```{r, include = TRUE, echo=TRUE}
var_data <- ts(var_data, start = start(inf$inflation), frequency = 12)
lagselect <- VARselect(var_data, lag.max = 12, type = "const")
lagselect
```

BIC (also known as Schwarz criterion) says to use `r lagselect$selection["SC(n)"]` lag(s). 

```{r, include = TRUE, echo = TRUE}
var_model <- VAR(var_data, p = lagselect$selection["SC(n)"], type = "const")
summary(var_model)
```

With a coefficient of **0.015**, significant at the 5% level, we see that lagged M4 growth positively affects inflation. For example, the mean M4 growth value of **`r round(mean(df_growth$money_growth_yoy, na.rm = TRUE,2))`%** would increase inflation by **`r round(0.015113 * mean(df_growth$money_growth_yoy, na.rm = TRUE),3)`** percentage points. The maximum M4 growth of **`r round(max(df_growth$money_growth_yoy, na.rm = TRUE,2))`%** would raise inflation by **`r round(0.015113 * max(df_growth$money_growth_yoy, na.rm = TRUE), 3)`** percentage points.

With a coefficient of **0.99** we also see that inflation is extremely dependent on lagged inflation, and therefore very persistent. 

Similarly, lagged inflation affects M4 growth with a coefficient of -0.05. So mean inflation growth of **`r round(mean(inf$inflation, na.rm = TRUE,2))`%** would decrease M4 growth by **`r round(mean(inf$inflation, na.rm = TRUE) * -0.04963, 2)`** percentage points. This captures monetary tightening by the Bank of England following an inflation spike. 

## Impulse response

Finally I plot an impulse-response function. This shows how inflation responds to a 1 S.D. shock to M4 growth, over a 25 month period. 

```{r, include = TRUE, echo=TRUE}
irf_model <- irf(var_model, impulse = "m4", response = "inflationn",
                 n.ahead = 24, boot = TRUE)
plot(irf_model)
```




