---
title: "Money Supply"
output: github_document
date: "2026-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(scales)
library(AER)
library(plm)
library(lmtest)
library(dplyr)
library(tidyr)
library(stargazer)
library(webshot2)
library(htmltools)
library(webshot2)
library(ggplot2)
library(tidyverse)
library(kableExtra)
```

```{r, include=FALSE}
source(here::here("money_supply.R"))
```

## Money Supply

Using data from <https://www.bankofengland.co.uk/statistics/Tables>, Table A2.2.1 - Components of M4, I plot the change in UK money supply. Since most of the money supply is created when commerical banks create deposits, I use total retail deposits for households and private non-financial corporations. Wholesale deposits and repos are excluded, as well as notes and coin. 

The data runs from `r min(df_growth$Date)` to `r max(df_growth$Date)`. I plot total money supply in sterling millions, month-on-month % growth rate as well as 3-month annualised growth rate. 

```{r plots, fig.height=2, fig.width=3, echo=FALSE, warning = FALSE}
ggplot(df, aes(x = Date, y = M4)) + 
  geom_line(color = "yellow", linewidth = 1.5) +
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Growth") +
  ggtitle("Money Supply, \n sterling millions 1982 - 2026") +
  theme(plot.title = element_text(size = 9 , face = "bold", colour = "white"))+
  theme(panel.background = element_rect("black")) +
  theme(plot.background = element_rect("black")) +
  theme(axis.text = element_text(colour = "white"))
```

The money supply as of `r format(max(df$Date))` is 
£`r scales::comma(1e6 * (df %>% filter(Date == max(Date)) %>% slice(1) %>% pull(M4)))`, i.e. £`r scales::comma(round((1e6 * (df %>% filter(Date == max(Date)) %>% slice(1) %>% pull(M4))) / 1e12,3))` trillion.






```{r more-plots, fig.height=2, fig.width=3, fig.show='hold', echo=FALSE, out.width="50%", echo=FALSE, warning=FALSE}
ggplot(df_growth, aes(x = Date, y = mom)) + 
  geom_line(color = "yellow", linewidth = 0.5) +
  labs("Year", "Growth") +
  scale_y_continuous(labels = comma) +
  ggtitle("% M4 growth - month-on-month") +
  theme(plot.title = element_text(size = 9 , face = "bold", colour = "white"))+
  theme(panel.background = element_rect("black")) +
  theme(plot.background = element_rect("black"))+
  theme(axis.text = element_text(colour = "white"))

ggplot(df_growth, aes(x = Date, y = growth_3m_ann)) + 
  geom_line(color = "yellow", linewidth = 0.5) +
  labs("Year", "Growth") +
  scale_y_continuous(labels = comma) +
  ggtitle("% M4 growth - 3-month annualised") +
  theme(plot.title = element_text(size = 9 , face = "bold", colour = "white"))+
  theme(panel.background = element_rect("black")) +
  theme(plot.background = element_rect("black"))+
  theme(axis.text = element_text(colour = "white"))
```

For month-on-month growth rates, the highest growth occurred on `r df_growth %>% filter(mom == max(mom, na.rm = TRUE)) %>% slice(1) %>% pull(Date)`, with a growth rate of `r paste0(round(df_growth %>% filter(mom == max(mom, na.rm = TRUE)) %>% slice(1) %>% pull(mom),2), "%")`.

For 3-month annualised growth rates, the highest growth occurred on `r df_growth %>% filter(growth_3m_ann == max(growth_3m_ann , na.rm = TRUE)) %>% slice(1) %>% pull(Date)`, with a growth rate of `r paste0(round(df_growth %>% filter(growth_3m_ann  == max(growth_3m_ann , na.rm = TRUE)) %>% slice(1) %>% pull(growth_3m_ann),2), "%")`.

## Inflation

I plot annualised inflation data from <https://www.ons.gov.uk/economy/inflationandpriceindices/timeseries/l55o/mm23>.

```{r inflation-plot, fig.height=2, fig.width=3, echo=FALSE}
ggplot(data = inflation, aes(x = Date, y = inflation)) +
  geom_line(color = "yellow", linewidth = 0.5) +
  labs(x = "Date", y = "Inflation") +
  ggtitle("Annual Inflation (%)") +
  theme(plot.title = element_text(face = "bold", size = 9))+
  theme(axis.title.x = element_text(color = "white"))+
  theme(axis.title.y = element_text(color = "white"))+
  theme(panel.background = element_rect("black")) +
  theme(plot.background = element_rect("black"))+
  theme(plot.title = element_text(color = "white"))
  
```

Between `r min(inflation$Date)` and `r max(inflation$Date)`, the highest recorded rate of inflation was **`r paste0(inflation %>% filter(inflation == max(inflation)) %>% slice(1) %>%  pull(inflation),"%")`** which occurred in `r inflation %>% filter(inflation == max(inflation)) %>% slice(1) %>% pull(Date)`